Vagueness, Noise, and Signaling

### 2015-09-10 paper/version_04/paper.tex
Created separate Discussion section. Added argument and citations for group selection.

This unifying property of perceptual imprecision could be considered an evolutionary beneficial
side-effect. Higher imprecision can lead to higher \emph{within group expected utility},
defined as the average expected utility that each evolved language scored when playing against
an arbitrary other language obtained for the same parameter
values. Figure~\ref{fig:GroupMeasuresB} gives a representative example. The observation repeats
for other parameter values: while imprecision might decrease the communicative efficiency of
individual languages, it increases the conceptual coherence and communicative success between
independently evolving strategies. It is as if mere confusability of states imposes a
regularity constraint on evolving categories.

@@

\section{Conclusion}
\label{sec:discussion}

We set out to meet a technical challenge posed by Lipman's problem
\citep{Lipman2009:Why-is-Language}: is there a conceptually sound and mathematically coherent
formal model that shows how vague language can evolve under selective pressure for efficient
communication? On the assumption that agents are likely to mistake states for one another,
especially when they are perceptually similar, derived a generalization of the replicator
dynamic in terms of imitation of successes. The resulting population-level dynamic produced
signaling behavior that is at the same time regular and by-and-large communicatively efficient,
while also showing the crucial marks of vagueness. In a sense, the model derives vagueness as a
by-product of an arguably natural limitation on the discriminatory power of signaling agents.

But the inclusion of confusability of similar states also had noteworthy effects on the
evolving meaning of signals. It transpired that imprecision can have further accelerating and,
surprisingly, unifying effects on meaning evolution. State confusability has an effect on
aggregate signaling behavior that can be described as generalization without generalizers: the
dynamics of imprecise imitation look as if ``conclusions'' about what works for one state are
``carried over'' to similar states. This, however, is merely an epiphenomenon in the sense no
single agent genuinely generalizes over stimuli or reasons about what a more systematic
signaling strategy would be.

In sum, inability to sharply discriminate similar stimuli may lead to vagueness and bring about
a (slight) decrease in communicative efficiency. But there may also be an advantage, not of
vagueness itself, but of its reason: systematicity in the confusability of states (which may be
a natural by-product of the perceptual system) supports \emph{as-if}-generalization at the
population-level without having to assume that agents themselves have any generalization power.

\citet{OConnor2013:The-Evolution-o} makes a related case based on a version of reinforcement
learning for sim-max games, in which successful play leads to reinforcement of choice options
also for states similar to the ones that actually occurred. This not only leads to vague
signaling of the appropriate kind, but also speeds up learning in such a way that, especially
for sim-max games with higher numbers of states, higher levels of communicative success are
reached in shorter learning periods. Our results complement and extend
\citeauthor{OConnor2013:The-Evolution-o}'s. The most important differences are that we obtain
similar regularizing effects also for cases with low numbers of states and do not assume that
agents have any kind of generalizing capacity in and of themselves, even if that is only
implicit, as in \citeauthor{OConnor2013:The-Evolution-o}'s generalized reinforcement learning.

@@

>>>>

@@

\section{Discussion}
\label{sec:discussion}

\subsection{Vagueness and evolution}
The inclusion of confusability of similar states has noteworthy effects on the
evolving meaning of signals. It transpired from our results that imprecision can have further accelerating and,
surprisingly, unifying effects on meaning evolution. State confusability has an effect on
aggregate signaling behavior that can be described as generalization without generalizers: the
dynamics of imprecise imitation look as if ``conclusions'' about what works for one state are
``carried over'' to similar states. This, however, is merely an epiphenomenon in the sense no
single agent genuinely generalizes over stimuli or reasons about what a more systematic
signaling strategy would be.

The unifying property of perceptual imprecision could be considered an evolutionary beneficial
side-effect. Higher imprecision can lead to higher \emph{within group expected utility},
defined as the average expected utility that each evolved language scored when playing against
an arbitrary other language obtained for the same parameter
values. Figure~\ref{fig:GroupMeasuresB} gives a representative example. The observation repeats
for other parameter values: while imprecision might decrease the communicative efficiency of
individual languages, it increases the conceptual coherence and communicative success between
independently evolving strategies. It is as if mere confusability of states imposes a
regularity constraint on evolving categories.

Although in our model the imprecision parameter was fixed for each simulation run, and thus not subject to evolutionary selection.
However, given the benefit of a certain amount of imprecision we observe when comparing within group expected utility, it would be interesting to study whether, under certain conditions, this group level advantage could trump the individual level disadvantage of a vague language, and thus actually select for a certain amount of imprecision.
This is for now purely speculative, but it makes sense especially if we consider it in the eyes of multilevel selection theory \citep{Wilson1994,OGorman2008}.

\subsection{Related work}
\citet{OConnor2013:The-Evolution-o} makes a related case based on a version of reinforcement
learning for sim-max games, in which successful play leads to reinforcement of choice options
also for states similar to the ones that actually occurred. This not only leads to vague
signaling of the appropriate kind, but also speeds up learning in such a way that, especially
for sim-max games with higher numbers of states, higher levels of communicative success are
reached in shorter learning periods. Our results complement and extend
\citeauthor{OConnor2013:The-Evolution-o}'s. The most important differences are that we obtain
similar regularizing effects also for cases with low numbers of states and do not assume that
agents have any kind of generalizing capacity in and of themselves, even if that is only
implicit, as in \citeauthor{OConnor2013:The-Evolution-o}'s generalized reinforcement learning.

@@

\section{Conclusion}
\label{sec:conclusion}

We set out to meet a technical challenge posed by Lipman's problem
\citep{Lipman2009:Why-is-Language}: is there a conceptually sound and mathematically coherent
formal model that shows how vague language can evolve under selective pressure for efficient
communication? On the assumption that agents are likely to mistake states for one another,
especially when they are perceptually similar, derived a generalization of the replicator
dynamic in terms of imitation of successes. The resulting population-level dynamic produced
signaling behavior that is at the same time regular and by-and-large communicatively efficient,
while also showing the crucial marks of vagueness. In a sense, the model derives vagueness as a
by-product of an arguably natural limitation on the discriminatory power of signaling agents.
Although inability to sharply discriminate similar stimuli may lead to vagueness and bring about
a (slight) decrease in communicative efficiency, there may also be an advantage, not of
vagueness itself, but of its reason: systematicity in the confusability of states (which may be
a natural by-product of the perceptual system) supports \emph{as-if}-generalization at the
population-level without having to assume that agents themselves have any generalization power.
